---
title: "p8105_hw2_yl4923"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1
1.1 Clean Mr.Trash Wheel data
```{r import and clean data}
trash_wheel = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N535") %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = trunc(sports_balls))
```

1.2 Clean and joint precipitation datasets
```{r binding precipitation}
precip_2018 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet = "2018 Precipitation",range = "A2:B14") %>%
  mutate(year = "2018") %>% 
  janitor::clean_names() %>%
  drop_na(total) 

precip_2019 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet = "2019 Precipitation",range = "A2:B14") %>%
  mutate(year = "2019") %>% 
  janitor::clean_names() %>%
  drop_na(total) 
   
precip_df = bind_rows(precip_2019, precip_2018) %>%
  mutate(month = month.name[month]) %>%
  select(year,month,total)
```
Trash Wheel data shows the types of trash and their corresponding amounts collected by dumpsters from _2014_ to _2020_, including `r nrow(trash_wheel)` observations. Such as for one of the trashes, sports balls, the median of it in tons collected in 2019 is *`r median(pull(filter(trash_wheel, year == 2019),sports_balls))`*. Over years, Mr. Trash Wheel collects *`r sum(trash_wheel$weight_tons)`* tons of trash.

For precipitation data from 2019 and 2018,we are able to identify the original source of the data by creating _year_ as a new variable.The precipitation data has `r nrow(precip_df)` rows and `r ncol(precip_df)` columns. The total precipitation in 2018 can be derived *`r sum(pull(filter(precip_df, year == 2018)))`*.

### Problem 2
2.1 Make 3 datasets jointable
```{r clean 3 datasets}
party = c("dem", "gop", "gop") 
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year","month","day")) %>%
  mutate(month = month.name[as.integer(month)], president = party[prez_gop + 1]) %>%
  select(year,month,president,everything(),-day, -prez_dem, -prez_gop)

snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(date, into = c("month","day","year")) %>%
  mutate(month = month.name[as.integer(month)]) %>%
  select(year,month,close,-day)

unemp_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemploy_percentage") %>%
  mutate(Year = as.character(Year)) %>%
  select(year = Year,month,unemploy_percentage)
```

2.2 Merging _snp_ into _pols_, and merging _unemployment_ into the result.
```{r Join 3 datasets}
pols_snp_unemploy = left_join(pols_df, snp_df, by = c("year", "month")) %>% 
  left_join(unemp_df, by = c("year", "month"))
```
The _pols_data_ has `r nrow(pols_df)` observations and `r ncol(pols_df)` variables, describing the party of president, numbers of government officials in a given year. The _snp_data_ contains `r nrow(snp_df)` observations and `r ncol(snp_df)` variables, showing the closing price of S&P overtime. The _unemployment data_ with `r nrow(unemp_df)` observations and `r ncol(unemp_df)` variables gives the percentage of unemployment in certain month and year. The joint of all three data sets by *year* and *month* has `r nrow(pols_snp_unemploy)` rows and `r ncol(pols_snp_unemploy)` columns from `r pull(pols_snp_unemploy, year)[1]` to `r rev(pull(pols_snp_unemploy, year))[1]`. We then can examine the politics, S&P stock market and unemployment altogether in a specific month and year.

